---
title: "Data Sci Boot Camp"
author: "Germán Silva"
date: "7/13/2022"
output: html_document
---

# Overview

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)

# significance tests and data exploration
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(ggbeeswarm)
library(broom)
library(ggfortify)
```

# Reading and Summarizing the data:

## Reading and cleaning data

### Bulk Density
```{r}
# pipe operator: Windows (ctrl+shift+m), MAC (cmd+shift+m)

soils <- read_csv(here("data", "tutorial_data.csv")) %>% # reads in the csv so we can work with it in the code
  clean_names() %>% 
  drop_na() %>%  # drops all rows with an NA 
  rename(ec = electro_cond_m_s_per_cm) %>% 
  mutate(water_content = ((wet_mass_g - dried_mass_g)/wet_mass_g) * 100)



soils_sum <- soils %>% # calls the averaged data
  group_by(transect) %>% #groups by zone only
  summarize(min = min(ec), # returns minimum value
            max = max(ec), # returns maximum value
            sd = sd(ec), # returns group's standard deviation
            mean = mean(ec), # returns group's mean
            sample_size = n()) # returns sample size

kable(soils_sum, # starts the creation of a table with the data from the object called
      col.names = c("Zone",
                    "Min. EC (mS/cm)",
                    "Max. EC (mS/cm)",
                    "Standard Deviation (mS/cm)",
                    "Mean EC (mS/cm)",
                    "Sample Size")) %>%  # renames all the column names to professional looking titles
  kable_classic(full_width = FALSE) # themes the table


```


## Visualizing

### Bulk Density
```{r}
ggplot(soils, aes(x = transect, y = ec, color = transect))+ 
  geom_beeswarm(aes(fill=transect), shape= 21, color = "black")+
  geom_boxplot(aes(fill=transect),
               color = "black",
               alpha=0.45,
               width = 0.2,
               outlier.color = NA)
```


```{r}
ggplot(soils, aes(x = elevation, y = ec, color = transect)) +#ggplot of soils/dem data
  geom_point()+#geometry of the plot
  labs(x = "Elevation (m)",#labels
       y = "Electrical conductivity (mS/cm)")

ggplot(soils, aes(x = elevation, y = ec, color = transect)) +#ggplot of soils/dem data
  geom_point()+#geometry of the plot
  facet_wrap(~transect)+
  labs(x = "Elevation (m)",#labels
       y = "Electrical conductivity (mS/cm)")
```
```{r}
ggplot(soils, aes(x = water_content, y = ec, color = transect)) +#ggplot of soils/dem data
  geom_point()+#geometry of the plot
  labs(x = "Elevation (m)",#labels
       y = "Electrical conductivity (mS/cm)")

ggplot(soils, aes(x = water_content, y = ec, color = transect)) +#ggplot of soils/dem data
  geom_point()+#geometry of the plot
  facet_wrap(~transect)+
  labs(x = "Water Content (%)",#labels
       y = "Electrical conductivity (mS/cm)")
```

# Hypothesis testing

Questions:

- Is there a significant difference in mean soil bulk density between the two zones? Is there a significant difference between longest branch length in the two zones?
- Does the data fit the assumptions of parametric significance testing?

## Exploring the data

### Soil Bulk Density

Histogram:
```{r}
ggplot(data = soils, aes(x = ec))+ # starts creation of graph, aes() refers to aesthetics and decides how things are displayed from the variables
  geom_histogram()+ # turns graph into a histogram
  facet_wrap(~location) # makes two panels based on Zone grouping

ggplot(data = soils, aes(x = ec))+ # starts creation of graph, aes() refers to aesthetics and decides how things are displayed from the variables
  geom_histogram()+ # turns graph into a histogram
  facet_wrap(~transect) # makes two panels based on Zone grouping
```


QQ plot:
```{r}
ggplot(soils, aes(sample = ec))+ # starts graph with mean bulk density as the sample/graphing data
  geom_qq()+ # makes a QQ plot, a exploratory/diagnostic plot
  facet_wrap(~location) # panels by Zone Grouping

ggplot(soils, aes(sample = ec))+ # starts graph with mean bulk density as the sample/graphing data
  geom_qq()+ # makes a QQ plot, a exploratory/diagnostic plot
  facet_wrap(~transect) # panels by Zone Grouping
```

## T-test

### Bulk Density

```{r}
copr <- soils %>% 
  filter(location == "COPR")

ncos <- soils %>% 
  filter(location == "NCOS")

cohen_d <- effsize::cohen.d(copr$ec, ncos$ec)

cohen_d

t_test <- t.test(copr$ec, ncos$ec)

t_test
```

# ANOVA

```{r}
soils$transect <- ordered(soils$transect,
                          levels = c('COPR_1_NS', 'COPR_2_EW', 'COPR_2_NS', 'NCOS_1_NS', 'NCOS_2_EW', 'NCOS_2_NS'))

soil_transect <- aov(ec ~ transect, data = soils)

summary(soil_transect)
```

# OLS and Multiple Linear Regression

```{r}
lm <- lm(ec ~ elevation,
   data = soils)

cor(x = soils$elevation, y = soils$ec)

summary(lm)

ggplot(soils, aes(x = elevation, y = ec)) +#ggplot of soils/dem data
  geom_point()+#geometry of the plot
  geom_smooth(method = 'lm')+
  labs(x = "Elevation (m)",#labels
       y = "Electrical conductivity (mS/cm)")
```  
  
```{r}
mlr <- glm(ec ~ location + water_content + landcover + elevation,
    data = soils)

summary(mlr)

```


